---
author: "Nicholas Reich"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
title: 'Summary of testing phase German flu forecasting results'
output: 
  pdf_document:
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(ForecastFramework)
library(kableExtra)
source("../code/forecast-utils.R")

theme_set(theme_bw())

testing_data <- readRDS("../data/testing_data.rds")
test_dat_long <- gather_data(testing_data)
```


```{r}
model_names <- tools::file_path_sans_ext(list.files(path = "../models", pattern = "*.R"))
model_names <- model_names[-which(model_names=="ContestModel")]
```



## Overview 

This document summarizes the performance of time-series forecasting models developed by UMass-Amherst Biostatistics PhD students for a PhD Seminar. 
The goal was to create forecasting models that accurately forecasted influenza incidence in Germany. We used as the gold-standard data a time series obtained from the Robert Koch Institut (RKI), a national public health agency in Germany. The testing dataset is shown below.

```{r test-dat, fig.cap="All data, stratified by state, with the three seasons evaluated for testing showed in darker color"}
ggplot(data=test_dat_long, aes(x=season_week, y=value, group=season)) +
    geom_line(alpha=.3) + ylab("incidence (per 100,000 population)") +
    geom_line(data=filter(test_dat_long, season %in% c("2016/2017", "2017/2018", "2018/2019"))) +
    facet_wrap(.~location) 
```


## Results across all regions
```{r, compile-all-results, results='asis'}
result_filenames <- list.files(path = "../results", pattern = "*testing-results.csv")
data_path <- "../results"
all_data <- tibble(filename = result_filenames) %>% # create a data frame
                                         # holding the file names
  mutate(file_contents = map(filename,          # read files into
           ~ read_csv(file.path(data_path, .))) # a new data column
        ) %>%
  unnest() %>%
  separate(filename, c("model"), sep="-")


data_forecasted <- all_data %>%
#data_forecasted <- fcast_data %>%
    left_join(dplyr::select(test_dat_long, location, date, value)) %>%
    rename(truth=value) %>%
    mutate(
        bias = pred_median - truth,
        ci50_cov = truth>pred_50_lb & truth<pred_50_ub,
        ci80_cov = truth>pred_80_lb & truth<pred_80_ub,
        ci95_cov = truth>pred_95_lb & truth<pred_95_ub) 

## overall summary
overall_summary <- data_forecasted %>% 
  dplyr::select(model, bias, starts_with("ci")) %>%
  group_by(model) %>%
  summarize(
    avg_bias = mean(bias, na.rm=TRUE),
    ci50_cov = mean(ci50_cov, na.rm=TRUE),
    ci80_cov = mean(ci80_cov, na.rm=TRUE),
    ci90_cov = mean(ci95_cov, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(model = reorder(model, abs(avg_bias))) %>%
  arrange(abs(avg_bias))
  
kable(overall_summary, "latex", booktabs = T, digits = 2) %>%
  kable_styling(latex_options = "striped")
```



## Average bias by region
```{r, regional-summary}
## overall summary
regional_bias_summary <- data_forecasted %>% 
  dplyr::select(model, location, bias, starts_with("ci")) %>%
  group_by(model, location) %>%
  summarize(avg_bias = mean(bias, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(model = reorder(model, abs(avg_bias))) %>%
  arrange(model) %>%
  spread(model, avg_bias)
  
kable(regional_bias_summary, "latex", booktabs = T, digits = 2) %>%
  kable_styling(latex_options = "striped")
```


## Average bias by target
```{r, target-summary}
## overall summary
target_bias_summary <- data_forecasted %>% 
  dplyr::select(model, step, bias, starts_with("ci")) %>%
  group_by(model, step) %>%
  summarize(avg_bias = mean(bias, na.rm=TRUE)) %>%
  ungroup() %>%
  mutate(model = reorder(model, abs(avg_bias))) %>%
  arrange(model) %>%
  spread(model, avg_bias)
  
kable(target_bias_summary, "latex", booktabs = T, digits = 2) %>%
  kable_styling(latex_options = "striped")
```

## some plots


```{r}
ggplot(data=filter(data_forecasted, timezero==as.Date("2019-02-24"), model=="DL4EPI"), aes(x=date)) +
    geom_line(data=test_dat_long, aes(y=value)) + 
    geom_point(data=test_dat_long, aes(y=value)) +
    geom_line(aes(y=pred_median), color="red") + 
    geom_point(aes(y=pred_median), color="red") + 
    geom_ribbon(aes(ymin = pred_50_lb, ymax = pred_50_ub), alpha = 0.2, fill="red") +
    geom_ribbon(aes(ymin = pred_80_lb, ymax = pred_80_ub), alpha = 0.2, fill="red") +
    geom_ribbon(aes(ymin = pred_95_lb, ymax = pred_95_ub), alpha = 0.2, fill="red") +
    facet_wrap(.~location) +
    scale_x_date(limits=as.Date(c("2018-10-01", "2019-06-01"))) +
    ggtitle("DL4EPI Forecasts from 2019 epiweek 9")
```

```{r}
ggplot(data=filter(data_forecasted, timezero==as.Date("2019-02-24"), model=="hetGP"), aes(x=date)) +
    geom_line(data=test_dat_long, aes(y=value)) + 
    geom_point(data=test_dat_long, aes(y=value)) +
    geom_line(aes(y=pred_median), color="red") + 
    geom_point(aes(y=pred_median), color="red") + 
    geom_ribbon(aes(ymin = pred_50_lb, ymax = pred_50_ub), alpha = 0.2, fill="red") +
    geom_ribbon(aes(ymin = pred_80_lb, ymax = pred_80_ub), alpha = 0.2, fill="red") +
    geom_ribbon(aes(ymin = pred_95_lb, ymax = pred_95_ub), alpha = 0.2, fill="red") +
    facet_wrap(.~location) +
    scale_x_date(limits=as.Date(c("2018-10-01", "2019-06-01"))) +
    ggtitle("hetGP Forecasts from 2019 epiweek 9")
```


```{r}
ggplot(data=filter(data_forecasted, timezero==as.Date("2019-02-24"), model=="SIRS_EAKF"), aes(x=date)) +
    geom_line(data=test_dat_long, aes(y=value)) + 
    geom_point(data=test_dat_long, aes(y=value)) +
    geom_line(aes(y=pred_median), color="red") + 
    geom_point(aes(y=pred_median), color="red") + 
    geom_ribbon(aes(ymin = pred_50_lb, ymax = pred_50_ub), alpha = 0.2, fill="red") +
    geom_ribbon(aes(ymin = pred_80_lb, ymax = pred_80_ub), alpha = 0.2, fill="red") +
    geom_ribbon(aes(ymin = pred_95_lb, ymax = pred_95_ub), alpha = 0.2, fill="red") +
    facet_wrap(.~location) +
    scale_x_date(limits=as.Date(c("2018-10-01", "2019-06-01"))) +
    ggtitle("SIRS_EAKF Forecasts from 2019 epiweek 9")
```

```{r}
ggplot(data=filter(data_forecasted, timezero==as.Date("2019-02-24"), model=="sarimaTD"), aes(x=date)) +
    geom_line(data=test_dat_long, aes(y=value)) + 
    geom_point(data=test_dat_long, aes(y=value)) +
    geom_line(aes(y=pred_median), color="red") + 
    geom_point(aes(y=pred_median), color="red") + 
    geom_ribbon(aes(ymin = pred_50_lb, ymax = pred_50_ub), alpha = 0.2, fill="red") +
    geom_ribbon(aes(ymin = pred_80_lb, ymax = pred_80_ub), alpha = 0.2, fill="red") +
    geom_ribbon(aes(ymin = pred_95_lb, ymax = pred_95_ub), alpha = 0.2, fill="red") +
    facet_wrap(.~location) +
    scale_x_date(limits=as.Date(c("2018-10-01", "2019-06-01"))) +
    ggtitle("sarimaTD Forecasts from 2019 epiweek 9")
```


