---
author: "Nicholas Reich"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
title: 'Summary of training phase forecasting results: German flu forecasting'
output: 
  pdf_document:
    fig_caption: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(ForecastFramework)
library(kableExtra)
source("../code/forecast-utils.R")

theme_set(theme_bw())

training_data <- readRDS("../data/training_data.rds")
train_dat_long <- gather_data(training_data)
```


```{r}
model_names <- tools::file_path_sans_ext(list.files(path = "../models", pattern = "*.R"))
model_names <- model_names[-which(model_names=="ContestModel")]
```



## Overview 

This document summarizes the performance of time-series forecasting models developed by UMass-Amherst Biostatistics PhD students for a PhD Seminar. 
The goal was to create forecasting models that accurately forecasted influenza incidence in Germany. We used as the gold-standard data a time series obtained from the Robert Koch Institut (RKI), a national public health agency in Germany. The training dataset is shown below.

```{r train-dat, fig.cap="Training data, stratified by state, with the one season evaluated for training showed in darker color", eval=FALSE}
ggplot(data=train_dat_long, aes(x=season_week, y=value, group=season)) +
    geom_line(alpha=.3) + ylab("incidence (per 100,000 population)") +
    geom_line(data=filter(train_dat_long, season=="2015/2016")) +
    facet_wrap(.~location) 
```



## Results across all regions
```{r, compile-all-results, results='asis'}
result_filenames <- list.files(path = "../results", pattern = "*training-results.csv")
data_path <- "../results"
all_data <- data_frame(filename = result_filenames) %>% # create a data frame
                                         # holding the file names
  mutate(file_contents = map(filename,          # read files into
           ~ read_csv(file.path(data_path, .))) # a new data column
        ) %>%
  unnest() %>%
  separate(filename, c("model"), sep="-")


data_forecasted <- all_data %>%
#data_forecasted <- fcast_data %>%
    left_join(dplyr::select(train_dat_long, location, date, value)) %>%
    rename(truth=value) %>%
    mutate(
        bias = pred_median - truth,
        ci50_cov = truth>pred_50_lb & truth<pred_50_ub,
        ci80_cov = truth>pred_80_lb & truth<pred_80_ub,
        ci95_cov = truth>pred_95_lb & truth<pred_95_ub) 

## overall summary
overall_summary <- data_forecasted %>% 
  dplyr::select(model, bias, starts_with("ci")) %>%
  group_by(model) %>%
  summarize(
    avg_bias = mean(bias, na.rm=TRUE),
    ci50_cov = mean(ci50_cov, na.rm=TRUE),
    ci80_cov = mean(ci80_cov, na.rm=TRUE),
    ci90_cov = mean(ci95_cov, na.rm=TRUE))
  
kable(overall_summary, "latex", booktabs = T, digits = 2) %>%
  kable_styling(latex_options = "striped")

```



## Results by region
```{r, regional-summary}
## overall summary
regional_bias_summary <- data_forecasted %>% 
  dplyr::select(model, location, bias, starts_with("ci")) %>%
  group_by(model, location) %>%
  summarize(avg_bias = mean(bias, na.rm=TRUE)) %>%
  spread(model, avg_bias)
  
kable(regional_bias_summary, "latex", booktabs = T, digits = 2) %>%
  kable_styling(latex_options = "striped")

```


